Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-09-12 ms.tex
forgot to save file
weird typo
Davids comments on v3

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were still sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. Using statistics and limits for the impact history on early Earth and the impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis (i.e., the origin of life) is favored early in the Hadean eon. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred in $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainty. Previous claims of a far narrower window for the origin of life are unsupported. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is possible in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for life on rocky planets elsewhere because they will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

\citet{Benner_2020} argued that the most likely time for the origin of life in an RNA-world scenario would have been $4.36 \pm 0.1$ Ga in the wake of a $\sim 2 \times 10^{22}$ kg ($\sim 2100$ km) asteroid impact called Moneta. They suggested that iron delivered by Moneta's core would have reacted with impact-vaporized ocean water to generate a reducing Hadean atmosphere conducive to the photochemical generation of essential prebiotic nitriles like HCN and HCCCN. Nitriles are required in prebiotic schemes that synthesize ribonucleotide precursors to RNA \citep{Patel_2015,Powner_2009,Becker_2019}. A transient period of profuse nitrile production after a Moneta impact could have been a window of opportunity for the origin of RNA and, ultimately, life.

A single Moneta impact could also explain the highly-siderophile elements (i.e. iron-loving, abbreviated HSEs) in the Moon's and the Earth's mantles. During the Moon-forming impact, HSEs should have been sequestered in each planetary core, so over-abundant HSEs in today's mantles are commonly interpreted as evidence for late-accretion impactors. Earth's mantle has substantially more HSEs than the Moon by an amount that cannot be accounted for by Earth's greater gravitational cross section \citep{Day_2015}. A single massive Moneta impact could explain the Earth-Moon HSE discrepancy because, by the statistics of small numbers, Moneta could have missed the Moon and hit the Earth \citep{Sleep_1989,Bottke_2010}.

@@

Recently, \citet{Wogan_2023} and \citet{Zahnle_2020} used photochemical models of post-impact atmospheres to show that impacts significantly smaller than Moneta can efficiently produce prebiotic molecules. In their simulations, impactor iron equilibrates with vaporized ocean water to generate atmospheric H$_2$, CH$_4$ and NH$_3$ that form thermochemically as the reducing steam atmosphere cools. Once steam condenses to an ocean, subsequent photochemistry of a Titan-like, hazy atmosphere produces prebiotic nitriles. \citet{Zahnle_2020} was a preliminary study that used simple atmospheric models, and later \citet{Wogan_2023} improved upon their calculations with more sophisticated and accurate simulations. Photochemical modeling in \citet{Wogan_2023} shows that the production of both HCN and HCCCN only occurs when $\mathrm{CH_4} / \mathrm{CO_2} \gtrsim 0.1$ while the atmosphere is hazy. In an optimistic scenario, which includes nickel-catalyzed methane production, results suggests that $\mathrm{CH_4} / \mathrm{CO_2} > 0.1$ (i.e. significant nitriles) occurs for impacts $> 4 \times 10^{20}$ kg ($> 570$ km diameter). In a pessimistic case, which assumes only a fraction of impactor iron reacts with the atmosphere \citep{Citron_2022}, a $> 5 \times 10^{21}$ kg ($> 1330$ km diameter) impactor is required to produce substantial HCN and HCCCN \citep{Wogan_2023}. Regardless of uncertainty, atmospheric models suggest that an impactor much smaller than Moneta could deliver the nitriles needed for an RNA origin of life.

Here, we use the \citet{Wogan_2023} results, along with Monte-Carlo simulations of Earth's impact history, to make an alternative estimate for when life most likely emerged during an RNA-first process. Our calculations account for the possibility of planet sterilization by impacts that vaporize the ocean \citep{Sleep_1989}. We assume that a life-starting impact is one that produces significant prebiotic nitriles and is not subsequently followed by an ocean-vaporizing impact that destroys the biosphere. By considering the fraction of stochastic impact realizations that do not have a life-starting impact, we also estimate the probability of life beginning if Earth history was rerun.

\section{Methods} \label{sec:methods}
The rate impacts hit Earth bigger than mass $m$ at age $t$ can be written

@@

We assume that the the majority of the size-frequency distribution of impactors is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. Data for the frequency of main-belt asteroids only extends down to about 1 km diameter ($1.3 \times 10^{12} $ kg). To extend the distribution to smaller objects, we use the observed 1400 ratio between the frequency of $> 1$ km and $> 20$ km craters on the Moon following \citet{Morbidelli_2018}. Crater scaling relations suggest that 1 km and 20 km craters on the Moon corresponds to 50 m and 1 km asteroids, respectively \citep{Morbidelli_2018}. The largest main belt asteroid is $\sim 1000$ km, so we must extrapolate to larger impactors. Through the main text, we use the same extrapolation as \citet{Marchi_2014}, which extends the distribution to big impacts with a slope $d (\ln S_0)/d (\ln m) = - 0.415$ (Figure \ref{fig:sfd_and_flux}). In Appendix \ref{sec:append_sfd} we show that instead extrapolating with the red dashed line in Figure \ref{fig:sfd_and_flux}a (with slope $d (\ln S_0)/d (\ln m) = - 1.0$) does not significantly change our overall conclusions. Finally, we normalize the size-frequency distribution to the impact mass required to make a 1 km crater on the moon (50 m object, $m_0 = 1.64 \times 10^{8}$ kg). Figure \ref{fig:sfd_and_flux}a shows the resulting size-frequency distribution.

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and, in the main text, log-log extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. In Appendix \ref{sec:append_sfd} we also consider the red dashed extrapolation to $\gtrsim 10^{21}$ kg impacts. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

The flux of impactors, $F_0(t)$, can be estimated from the lunar cratering record. We adopt the accretionary tail scenario discussed in \citet{Morbidelli_2018}. In other words, we assume that the ``late heavy bombardment'' did not happen \citep{Cartwright_2022,Hartmann_2019,Zellner_2017}, and that the impact flux on Earth was monotonically decreasing throughout the Hadean eon. Figure \ref{fig:sfd_and_flux}b shows $n_m$, our assumed lunar impact history taken from \citet{Morbidelli_2018} (red line in their Figure 5). We use the parameterization $n_m = 10^{a e^{b t} + c}$, where $a$, $b$, and $c$ are fit parameters given in Figure \ref{fig:sfd_and_flux}b.

Extrapolating the lunar impact history to Earth requires correcting for Earth's greater gravitational attraction. Assuming that the approach velocity of impactors far from the Earth and Moon was on average 18 km/s \citep{Morbidelli_2018}, then Earth should receive $s_0 = 1.36$ times more impacts per surface area than the Moon. Therefore, the number of impacts on Earth is

\begin{equation}
  N_0 = A_\oplus s_0 n_m = A_\oplus s_0 10^{a e^{b t} + c}
  \label{eq:num_earth_impacts}
\end{equation}
Equation \eqref{eq:num_earth_impacts} also accounts for Earth's surface area ($A_\oplus$), making $N_0$ the number of impacts on Earth since age $t$ that would cause a crater $> 1$ km on the moon. As discussed previously, \citet{Morbidelli_2018} use crater scaling relations to show that a 1 km lunar crater corresponds to a 50 m object with mass $m_0 = 1.64 \times 10^8$ kg.

@@

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Appendix Figure \ref{fig:velocity_distribution} shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close-approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10 km$). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/example_impact_histories.pdf}
  \caption{Three simulated impact histories of the 5000 derived from our Monte-Carlo approach (Section \ref{sec:methods}). Each vertical line indicates an impact of a mass shown by the y-axis. The red lines indicate the last impact to vaporize the ocean assuming 10\% of an impacts kinetic energy heats the ocean. (a) gives conversions between impact mass and diameter for 500, 1000 and 2000 km asteroids.}
  \label{fig:example_impact_histories}
\end{figure}

Figure \ref{fig:example_impact_histories} illustrates three of the 5000 impact histories computed with our Monte-Carlo approach (Section \ref{sec:methods}). In Figure \ref{fig:example_impact_histories}a, a Moneta-sized impact occurs at $\sim 4.43$ Ga delivering nearly all of Earth's mantle HSEs. The impact should also reduce the Hadean atmosphere creating conditions favorable for prebiotic chemistry. However, in this case, Moneta is not a life-starting impact because Moneta is followed by an ocean-vaporizing collision at $\sim 4.37$ Ga which we assume destroys any existing biosphere. Based on smoothed-particle hydrodynamics impact simulations, we nominally assume that 10\% of an impact's kinetic energy goes into vaporizing the ocean \citep[Appendix \ref{sec:append_vap},][]{Citron_2022}. Vaporizing the ocean requires $5 \times 10^{27}$ J, so a $> 5 \times 10^{28}$ J impact should deliver sufficient energy under our assumptions. Appendix \ref{sec:append_vap} considers alternative criteria for ocean-vaporization. In Figure \ref{fig:example_impact_histories}a, the last ocean-vaporizing impact is only $\sim 10^{20}$ kg (360 km) but has a large impact velocity of $32$ km/s giving it $5.12 \times 10^{28}$ J of kinetic energy.

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would cause biopoiesis depends on the required minimum impact mass to produce significant origin-of-life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} uses photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} finds that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts.pdf}
  \caption{The probability of an impact that causes favorable conditions for an origin of life on Earth. The blue solid line is the probability of an impact occurring at least once. The orange dashed line is the probability of an impact without a subsequent ocean-vaporizing collision. Probabilities are shown as a function of the minimum impact mass to produce significant prebiotic molecules (e.g. HCN). Two plausible minimum masses are the \citet{Wogan_2023} optimistic and pessimistic scenarios indicated with vertical dotted lines. The shaded region between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg are Moneta-sized impacts because they deliver most all of Earth's mantle HSEs. There is a 92\% chance of an impact creating favorable origin of life conditions in the \citet{Wogan_2023} optimistic case, and a 41\% chance in the pessimistic case.}
  \label{fig:probabilities_of_impacts}
\end{figure}

As illustrated in Figure \ref{fig:example_impact_histories}, a life-starting impact does not occur in every simulated impact history. In our impact-driven model for the origin of life, biopoiesis requires (1) an impact of sufficient mass to make prebiotic nitriles and (2) the lack of a later ocean-vaporizing collision that destroys the biosphere without rekindling it. Figure \ref{fig:probabilities_of_impacts} shows the probability of both conditions occurring as a function of the minimum impact mass that produces prebiotic molecules. For example, consider the \citet{Wogan_2023} optimistic minimum mass of $4 \times 10^{20}$ kg indicated in Figure \ref{fig:probabilities_of_impacts}. All simulated impact histories have at least one impact bigger than $4 \times 10^{20}$ kg (blue line in Figure \ref{fig:probabilities_of_impacts} is 1.0 at the given mass). However, the probability of the origin of life in this scenario is 92\%, because 8\% of the time the last $4 \times 10^{20}$ kg collision is followed by a smaller ocean-vaporizing impact that perhaps sterilizes the planet (orange dashed line in Figure \ref{fig:probabilities_of_impacts} with 0.92 at $4 \times 10^{20}$ kg). An origin of life is less probable if we consider the \citet{Wogan_2023} pessimistic minimum mass ($> 5 \times 10^{21}$ kg) to produce significant origin-of-life nitriles. A $> 5 \times 10^{21}$ kg impact without later ocean vaporization occurs 41\% of the time. A life-starting impact was therefore relatively likely on the early Earth, and was strongly favored (92\% probability) under optimistic assumptions.

@@

The timing of an origin of life impact is qualitatively unchanged when instead adopting the \citet{Wogan_2023} pessimistic case, which requires a $> 5 \times 10^{21}$ kg impact without later ocean vaporization (Figure \ref{fig:time_and_mass}b). In this scenario, the median timing for life's origin is 4.36 Ga with a 95\% confidence interval spanning 3.94 to 4.49 Ga (Figure \ref{fig:time_and_mass}b (i)). Under these assumptions, the life-starting impact is Moneta-sized 31\% of the time (Figure \ref{fig:time_and_mass}b (ii)), which is more probable compared to ``optimistic'' case (Figure \ref{fig:time_and_mass}a (ii)).

@@

Our estimated timing for biopoiesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology \citep{Rosing_1999,Ohtomo_2014}. Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be generally compatible with our estimated timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga.

Furthermore, Hadean zircons provide some constraint on Earth's impact history which we do not explicitly account for in our calculations. The oldest zircons are from Jack Hills, Australia, with U-Pb dates as old as $\sim 4.38$ Ga \citep{Valley_2014}. \citet{Benner_2020} argued that a Moneta-scale impact ($\sim 2 \times 10^{22}$ kg) could not have occurred after this date, because such a collision would heat the crust enough to potentially reset all U-Pb chronometers. \citet{Benner_2020} makes this claim based on \citet{Abramov_2013}, who used models of impact ejecta coupled to simulations of radiogenic Pb-loss in zircons to investigate chronometers resetting during the hypothesized ``late-heavy bombardment''. However, \citet{Abramov_2013} considers impacts much smaller than Moneta (e.g. $< 10^{21}$ kg), and does not clearly establish a minimum impactor mass that would reset all zircon chronometers globally. Given this ambiguity, our simulations do not use zircon chronometers as a constraint on Earth's impact history.

>>>>

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were still sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. We assume that prebiotic chemistry required a post-impact reducing atmosphere. Then, using statistics for the impact history on early Earth and the minimum impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis (i.e., the origin of life) is favored early in the Hadean eon. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred in a $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainty. Previous claims of a far narrower window for the origin of life are unsupported. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is possible in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for biopoesis on rocky planets elsewhere which will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

\citet{Benner_2020} argued that the most likely time for the origin of life would have been $4.36 \pm 0.1$ Ga in the wake of a $\sim 2 \times 10^{22}$ kg ($\sim 2100$ km diameter) asteroid impact called Moneta. They suggested that iron delivered by Moneta's core would have reacted with impact-vaporized ocean water to generate a reducing Hadean atmosphere conducive to the photochemical generation of essential prebiotic nitriles like HCN and HCCCN although their production rates were not modeled. Nitriles are required in prebiotic schemes that synthesize ribonucleotide precursors to RNA \citep{Patel_2015,Powner_2009,Becker_2019}. A transient period of profuse nitrile production after a Moneta impact could have been a window of opportunity for the origin of RNA and, ultimately, life.

A single Moneta impact could also explain a relative excess of highly-siderophile elements (i.e. iron-loving, abbreviated HSEs) in the Moon's and the Earth's mantles. During the Moon-forming impact, HSEs should have been sequestered in each planetary core, so over-abundant HSEs in today's mantles are commonly interpreted as evidence for late-accretion impactors. Earth's mantle has substantially more HSEs than the Moon by an amount that cannot be accounted for by Earth's greater gravitational cross section \citep{Day_2015}. A single massive Moneta impact could explain the Earth-Moon HSE discrepancy because, by the statistics of small numbers, Moneta could have missed the Moon and hit the Earth \citep{Sleep_1989,Bottke_2010}.

@@

Recently, \citet{Wogan_2023} and \citet{Zahnle_2020} used photochemical models of post-impact atmospheres to show that impacts significantly smaller than Moneta can efficiently produce molecules critical for prebiotic synthesis. In their simulations, impactor iron equilibrates with vaporized ocean water to generate atmospheric H$_2$, CH$_4$ and NH$_3$ that form thermochemically as the reducing steam atmosphere cools. Once steam condenses to an ocean, subsequent photochemistry of a Titan-like, hazy atmosphere produces prebiotic nitriles.

\citet{Zahnle_2020} was a preliminary study that used simple atmospheric models, and later \citet{Wogan_2023} improved upon their calculations with more sophisticated and accurate simulations. Photochemical modeling in \citet{Wogan_2023} shows that the production of both HCN and HCCCN only occurs when $\mathrm{CH_4} / \mathrm{CO_2} \gtrsim 0.1$ while the atmosphere is hazy. In an optimistic scenario, which includes nickel-catalyzed methane production, results suggests that $\mathrm{CH_4} / \mathrm{CO_2} > 0.1$ (i.e. significant nitriles) occurs for impacts $> 4 \times 10^{20}$ kg ($> 570$ km diameter). In a pessimistic case, which assumes only a fraction of impactor iron reacts with the atmosphere \citep{Citron_2022}, a $> 5 \times 10^{21}$ kg ($> 1330$ km diameter) impactor is required to produce substantial HCN and HCCCN \citep{Wogan_2023}. Both size limits may even be conservative, given experimental results showing that additional catalysts can generate methane \citep{Hill_2003,Nuth_2016,Nuth_2020}. Regardless of uncertainty, atmospheric models suggest that an impactor much smaller than Moneta could deliver the nitriles needed for an RNA origin of life.

Here, we use the \citet{Wogan_2023} results, along with Monte-Carlo simulations of Earth's impact history, to make an alternative estimate for when life most likely emerged during an RNA-first process. Our calculations account for the possibility of planet sterilization by impacts that vaporize the ocean \citep{Sleep_1989}. We assume that a life-starting impact is one that produces significant prebiotic nitriles and is not subsequently followed by an ocean-vaporizing impact that destroys any biosphere that has emerged. By considering the fraction of stochastic impact realizations that do not have a life-starting impact, we also estimate the probability of life beginning if Earth history was rerun.

\section{Methods} \label{sec:methods}
The rate impacts hit Earth bigger than mass $m$ (in kg) at age $t$ (in Gyr) can be written

@@

We assume that the the majority of the size-frequency distribution of impactors is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. Data for the frequency of main-belt asteroids only extends down to about 1 km diameter ($1.3 \times 10^{12} $ kg). To extend the distribution to smaller objects, we use the observed 1400 ratio between the frequency of $> 1$ km and $> 20$ km craters on the Moon following \citet{Morbidelli_2018}. Crater scaling relations suggest that 1 km and 20 km craters on the Moon corresponds to 50 m and 1 km asteroids, respectively \citep{Morbidelli_2018}. The largest main belt asteroid is $\sim 1000$ km, so we must extrapolate to larger impactors. Through the main text, we use the same extrapolation as \citet{Marchi_2014}, which extends the distribution to big impacts with a slope $d (\ln S_0)/d (\ln m) = - 0.415$ (Figure \ref{fig:sfd_and_flux}). In Appendix \ref{sec:append_sfd}, we show that alternatively extrapolating with the red dashed line in Figure \ref{fig:sfd_and_flux}a (with slope $d (\ln S_0)/d (\ln m) = - 1.0$) does not significantly change our overall conclusions. Finally, we normalize the size-frequency distribution to the impact mass required to make a 1 km crater on the Moon (50 m object, $m_0 = 1.64 \times 10^{8}$ kg). Figure \ref{fig:sfd_and_flux}a shows the resulting size-frequency distribution.

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution (SFD) of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and, in the main text, log-log extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. In Appendix \ref{sec:append_sfd} we also consider the red dashed extrapolation to $\gtrsim 10^{21}$ kg impacts. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

The flux of impactors, $F_0(t)$, can be estimated from the lunar cratering record. We adopt the accretionary tail scenario discussed in \citet{Morbidelli_2018}. In other words, we assume that the ``late heavy bombardment'' did not happen \citep{Cartwright_2022,Hartmann_2019,Zellner_2017}, and that the impact flux on Earth was monotonically decreasing throughout the Hadean eon. Figure \ref{fig:sfd_and_flux}b shows $n_m$, our assumed lunar impact history taken from \citet{Morbidelli_2018} (red line in their Figure 5). We use the parameterization $n_m = 10^{a e^{b t} + c}$, where $a$, $b$, and $c$ are parameter fits given in Figure \ref{fig:sfd_and_flux}b.

Extrapolating the lunar impact history to Earth requires correcting for Earth's greater gravitational attraction. Assuming that the approach velocity of impactors far from the Earth and Moon was on average 18 km/s \citep{Morbidelli_2018}, using Equation (7-14) in \citet{Zahnle_2006}, we compute that Earth should receive $s_0 = 1.36$ times more impacts per surface area than the Moon. Therefore, the number of impacts on Earth is

\begin{equation}
  N_0 = A_\oplus s_0 n_m = A_\oplus s_0 10^{a e^{b t} + c}
  \label{eq:num_earth_impacts}
\end{equation}
Equation \eqref{eq:num_earth_impacts} also accounts for Earth's surface area ($A_\oplus$), making $N_0$ the number of impacts on Earth since age $t$ that would cause a crater $> 1$ km on the Moon. As discussed previously, \citet{Morbidelli_2018} use crater scaling relations to show that a 1 km lunar crater corresponds to a 50 m object with mass $m_0 = 1.64 \times 10^8$ kg.

@@

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Appendix Figure \ref{fig:velocity_distribution} shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close-approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10$ km). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/example_impact_histories.pdf}
  \caption{Three simulated impact histories of the 5000 derived from our Monte-Carlo approach (Section \ref{sec:methods}). Each vertical line indicates an impact of a mass shown by the y-axis. The red lines indicate the last impact to vaporize the ocean assuming 10\% of an impactor's kinetic energy heats the ocean. The legend in (a) gives conversions between impact mass and diameter for 500, 1000 and 2000 km asteroids.}
  \label{fig:example_impact_histories}
\end{figure}

Figure \ref{fig:example_impact_histories} illustrates three of 5000 impact histories computed with our Monte-Carlo approach (Section \ref{sec:methods}). In Figure \ref{fig:example_impact_histories}a, a Moneta-sized impact occurs at $\sim 4.43$ Ga delivering nearly all of Earth's mantle HSEs. The impact should also chemically reduce the Hadean atmosphere creating conditions favorable for prebiotic chemistry. However, in this case, Moneta cannot have given rise to extant life on Earth because Moneta is followed by an ocean-vaporizing collision at $\sim 4.37$ Ga which should destroy any existing biosphere. Based on smoothed-particle hydrodynamics impact simulations, we nominally assume that 10\% of an impact's kinetic energy goes into vaporizing the ocean \citep[Appendix \ref{sec:append_vap},][]{Citron_2022}. Vaporizing the ocean requires $5 \times 10^{27}$ J, so a $> 5 \times 10^{28}$ J impact should deliver sufficient energy under our assumptions. Appendix \ref{sec:append_vap} considers alternative criteria for ocean-vaporization. In Figure \ref{fig:example_impact_histories}a, the last ocean-vaporizing impact is only $\sim 10^{20}$ kg (360 km) but has a large impact velocity of $32$ km/s giving it $5.12 \times 10^{28}$ J of kinetic energy.

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would cause biopoiesis depends on the required minimum impact mass to produce significant origin-of-life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} used photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} found that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts.pdf}
  \caption{The probability of an impact that causes favorable conditions for an origin of life on Earth. The blue solid line is the probability of an impact occurring at least once. The orange dashed line is the probability of an impact without a subsequent ocean-vaporizing collision. Probabilities are shown as a function of the minimum impact mass to produce significant prebiotic molecules (e.g. HCN). Two plausible minimum masses are the \citet{Wogan_2023} optimistic and pessimistic scenarios indicated with vertical dotted lines. The shaded region between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg are Moneta-sized impacts because they deliver most all of Earth's mantle HSEs. There is a 92\% chance of an impact creating favorable origin of life conditions in the \citet{Wogan_2023} optimistic case, and a 41\% chance in the pessimistic case. The chance of a Moneta-size impact giving rise to life that is not subsequently destroyed is 12\%.}
  \label{fig:probabilities_of_impacts}
\end{figure}

As illustrated in Figure \ref{fig:example_impact_histories}, a life-starting impact does not occur in every simulated impact history. In our impact-driven model for the origin of life, biopoiesis requires (1) an impact of sufficient mass to make prebiotic nitriles and (2) the lack of a later ocean-vaporizing collision that destroys the biosphere without rekindling it. Figure \ref{fig:probabilities_of_impacts} shows the probability of both conditions occurring as a function of the minimum impact mass that produces prebiotic molecules.

For example, consider the \citet{Wogan_2023} optimistic minimum mass of $4 \times 10^{20}$ kg indicated in Figure \ref{fig:probabilities_of_impacts}. All simulated impact histories have at least one impact bigger than $4 \times 10^{20}$ kg (blue line in Figure \ref{fig:probabilities_of_impacts} is 1.0 at the given mass). However, the probability of the origin of life in this scenario is 92\%, because 8\% of the time the last $4 \times 10^{20}$ kg collision is followed by a smaller ocean-vaporizing impact that perhaps sterilizes the planet (orange dashed line in Figure \ref{fig:probabilities_of_impacts} with 0.92 at $4 \times 10^{20}$ kg). An origin of life is less probable if we consider the \citet{Wogan_2023} pessimistic minimum mass ($> 5 \times 10^{21}$ kg) to produce significant origin-of-life nitriles. A $> 5 \times 10^{21}$ kg impact without later ocean vaporization occurs 41\% of the time. A life-starting impact was therefore relatively likely on the early Earth, and was strongly favored (92\% probability) under optimistic assumptions.

@@

The timing of an origin of life impact is qualitatively unchanged when instead adopting the \citet{Wogan_2023} pessimistic case, which requires a $> 5 \times 10^{21}$ kg impact without later ocean vaporization (Figure \ref{fig:time_and_mass}b). In this scenario, the median timing for life's origin is 4.36 Ga with a 95\% confidence interval spanning 3.94 to 4.49 Ga (Figure \ref{fig:time_and_mass}b (i)). Under these assumptions, the life-starting impact is Moneta-sized 31\% of the time with no subsequent sterilization (Figure \ref{fig:time_and_mass}b (ii)), which is more probable compared to ``optimistic'' case (Figure \ref{fig:time_and_mass}a (ii)).

@@

Our estimated timing for biopoiesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology \citep{Rosing_1999,Ohtomo_2014}. Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be generally compatible with our estimated timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga. In general, if life arose from post-impact reducing atmospheres, the origin of life was Hadean, not Archean.

Hadean zircons provide some constraint on Earth's impact history which we do not explicitly account for in our calculations. The oldest zircons are from Jack Hills, Australia, with U-Pb dates as old as $\sim 4.38$ Ga \citep{Valley_2014}. \citet{Benner_2020} argued that a Moneta-scale impact ($\sim 2 \times 10^{22}$ kg) could not have occurred after this date, because such a collision would heat the crust enough to potentially reset all U-Pb chronometers. \citet{Benner_2020} makes this claim based on \citet{Abramov_2013}, who used models of impact ejecta coupled to simulations of radiogenic Pb-loss in zircons to investigate chronometers resetting during the hypothesized ``late-heavy bombardment''. However, \citet{Abramov_2013} considers impacts much smaller than Moneta (e.g. $< 10^{21}$ kg), and does not clearly establish a minimum impactor mass that would reset all zircon chronometers globally. Given this ambiguity, our simulations do not use zircon chronometers as a constraint on Earth's impact history.

Even if an impact failed to reset a zircon chronometer, the shock wave produced by an asteroid collision may create micro- to nano-structural features in zircons that could be preserved over billions of years \citep{Reimink_2023}. Therefore, the presence or absence of shocked Hadean zircons in the geologic record may provide some constraint on Earth's impact history. For example, \citet{Reimink_2023} estimated the probability of preserved zircons with shock features assuming Earth experienced a ``late-heavy bombardment'' at 3.9 Ga. In this scenario, they find that shocked zircons were likely preserved yet find none in a collection of 4.02 Ga zircons from the Acasta Gneiss Complex. Overall, the absence of preserved shocked zircons suggests that a ``late-heavy bombardment'' did not occur at 3.9 Ga. Our Monte-Carlo models of Earth's impact history do not use zircon shock features as a constraint. A better model would incorporate this information, noting that resolving the issue of preservational bias of zircons is beyond the focus and scope of this paper.
